---
title: "Exploratory Data plotting Project2"
author: "SEH"
date: "10/11/2020"
output: html_document
---

```{r }
library(plyr)
library(ggplot2)
library(grid)
library(data.table)
library(dplyr)

## Data download

fileUrl <- "https://d396qusza40orc.cloudfront.net/exdata%2Fdata%2FNEI_data.zip"
download.file(fileUrl, destfile="~/coursera_ref/test/ExDataPlotting_project2/data.zip")
unzip("data.zip")

NEI<- readRDS("summarySCC_PM25.rds")
head(NEI)

 
SCC <- readRDS("Source_Classification_Code.rds")
head(SCC)

```

## Question1

Have total emissions from PM2.5 decreased in the United States from 1999 to 2008? Using the base plotting system, make a plot showing the total PM2.5 emission from all sources for each of the years 1999, 2002, 2005, and 2008.

```{r }
## question 1 
totpm25 <- aggregate(Emissions ~ year, NEI, FUN=sum)
barplot(height=totpm25$Emissions/1000000, names.arg=totpm25$year, xlab="Year", ylab=expression('Total PM2.5 Emission (x 10^6 tons)'))

png("plot1.png") ## plot1
barplot(height=totpm25$Emissions/1000000, names.arg=totpm25$year, xlab="Year", ylab=expression('Total PM2.5 Emission (x 10^6 tons)'))
dev.off() 
```

## Question2

Have total emissions from PM2.5 decreased in the Baltimore City, Maryland (\color{red}{\verb|fips == "24510"|}fips == "24510") from 1999 to 2008? Use the base plotting system to make a plot answering this question.

```{r }
Baltimore <- subset (NEI, fips == "24510")
totpm25Bal <- aggregate(Emissions ~ year, Baltimore, FUN=sum)
head(totpm25Bal)
barplot(height=totpm25Bal$Emissions/1000000, names.arg=totpm25Bal$year, xlab="Year", ylab=expression('Total PM2.5 Emission (x 10^6 tons)'))

png("plot2.png")
barplot(height=totpm25Bal$Emissions/1000000, names.arg=totpm25Bal$year, xlab="Year", ylab=expression('Total PM2.5 Emission (x 10^6 tons)'))
dev.off()

```

##Question3

Of the four types of sources indicated by the \color{red}{\verb|type|}type (point, nonpoint, onroad, nonroad) variable, which of these four sources have seen decreases in emissions from 1999–2008 for Baltimore City? Which have seen increases in emissions from 1999–2008? Use the ggplot2 plotting system to make a plot answer this question.

```{r }
Baltimore <- subset(NEI, fips == "24510")
typetotpm25Bal <- ddply(Baltimore, .(year, type), function(x) sum(x$Emissions))
colnames(typetotpm25Bal)[3] <- "Emissions"
p <- qplot(year, Emissions, data=typetotpm25Bal, color=type, geom ="line") + xlab("Baltimore(year)") + ylab("Total PM2.5 Emissions (in tons)")

png("plot3.png")
print(p)
dev.off()

```
### Answer: 'increased ...type(point) in 2005 and 'decreased ...type('non-road', 'nonpoint' and 'on-road')  
 
##Question4

Across the United States, how have emissions from coal combustion-related sources changed from 1999–2008?

```{r }
## question 4

combustionRelated <- grepl("comb", SCC$SCC.Level.One, ignore.case=TRUE)
coalRelated <- grepl("coal", SCC$SCC.Level.Four, ignore.case=TRUE) 
coalCombustion <- (combustionRelated & coalRelated)
combustionSCC <- SCC[coalCombustion,]$SCC
combustionNEI <- NEI[NEI$SCC %in% combustionSCC,]

png("plot4.png")
coalcomb <- aggregate(Emissions ~ year, combustionNEI, FUN=sum)
p <- qplot(year, Emissions, data=coalcomb, geom ="line") + xlab("coal_combustions(year)") + ylab("Total PM2.5 Emissions (in tons)")
print(p)
dev.off()
```
### answer: overall Emissions from coal combustion has been decreased from 2002. 

##Question5

How have emissions from motor vehicle sources changed from 1999–2008 in Baltimore City?
```{r }
## question5 
head(SCC)
vehiclesrelated <- grepl("vehicle", SCC$SCC.Level.Two, ignore.case=TRUE)
vehiclesSCC <- SCC[vehiclesrelated,]$SCC
vehiclesNEI <- NEI[NEI$SCC %in% vehiclesSCC,]
VehiclesNEIBal <- vehiclesNEI[vehiclesNEI$fips=="24510",]

vehicleBal<- aggregate(Emissions ~ year, VehiclesNEIBal, FUN=sum)
png("plot5.png")
p <- qplot(year, Emissions, data=vehicleBal, geom ="line") + xlab("Motor_vehicles(year)") + ylab("Total PM2.5 Emissions (in tons)")
print(p)
dev.off()
```
### Answer: Motor vehicle sourced emission in Baltomore has been significantly decreased from 2002. 


## Question6
Compare emissions from motor vehicle sources in Baltimore City with emissions from motor vehicle sources in Los Angeles County, California (\color{red}{\verb|fips == "06037"|}fips == "06037"). Which city has seen greater changes over time in motor vehicle emissions?

```{r }
## question 6
vehiclesrelated <- grepl("vehicle", SCC$SCC.Level.Two, ignore.case=TRUE)
vehiclesSCC <- SCC[vehiclesrelated,]$SCC
vehiclesNEI <- NEI[NEI$SCC %in% vehiclesSCC,]

VehiclesNEIBal <- vehiclesNEI[vehiclesNEI$fips=="24510",]
VehiclesNEILA <- vehiclesNEI[vehiclesNEI$fips=="06037",]


vehicleEmission_city <- ddply(VehiclesNEIBal_LA, .(year, fips), function(x) sum(x$Emissions))
colnames(vehicleEmission_city)[3] <- "Emissions"
VehiclesNEIBal$city <- "Baltimore"
VehiclesNEILA$city <- "Los Angeles"
VehiclesNEIBal_LA <- rbind(VehiclesNEIBal, VehiclesNEILA)
dim(VehiclesNEIBal)
dim(VehiclesNEILA)
dim(VehiclesNEIBal_LA)

png("plot6.png")
ggplot(VehiclesNEIBal_LA, aes(x=factor(year), y=Emissions)) +
  geom_bar(stat="identity") +
  facet_grid(scales="free", .~city) +
  xlab("year")+ ylab("Total PM2.5 Emissions (in tons)")
dev.off()
```
### Answer: Baltimore